---
title: "Clase 8: Supervivencia de nido"
author: "David Murillo"
format:
  html:
    css: styles.css
---

## Supervivencia de nidos

Para este analisi necesitaremos instalar el software [Jags](https://sourceforge.net/projects/mcmc-jags/)


## cargar paquetes

```{r}

library(jagsUI)
library(tidyverse)
```

## Cargar base de datos

```{r}
NidosPatos <- read.csv("data/duck_nests.csv", 
                   header=TRUE)
```

La base de datos duck_NidoPatos1s.csv contiene una muestra aleatoria de registro para tres años, divididos en 12 bloques. 

Primero estandarizaremos la visita inicial a cada nido

```{r}
NidoPatos1 <- NidosPatos %>% 
  mutate(init_stadard = (init- mean(init))/sd(init))
```

## Preparar datos para el modelo

Crearemos una lista con las covariables que pueden implementarce en el modelo

```{r}

dsr.data <- list(n.nests = dim(NidoPatos1)[1], 
                 n.yrs = max(NidoPatos1$year), 
                 n.areas = max(NidoPatos1$block), 
                 n.spp = max(NidoPatos1$species),
                 year = NidoPatos1$year, 
                 area = NidoPatos1$block, 
                 spp = NidoPatos1$species,
                 expos = NidoPatos1$expos, 
                 init = NidoPatos1$init_stadard,
                 age = NidoPatos1$age_start, 
                 survive = NidoPatos1$survive,
                 dist = NidoPatos1$dist)
```

## Modelo Nulo

```{r}
sink("dsr_null.jags")
cat("
model {
  # modeling daily nest survival on the logit scale
  dsr.mu ~ dnorm(0, 0.33) # intercept, vague prior (0,1) on real scale

  for (i in 1:n.nests) {
    for (j in 1:expos[i]) {
      # daily survival is a constant mean survival (no covariates or random effects)
      logitS[i,j] <- dsr.mu
      logit(S[i,j]) <- logitS[i,j] # convert back to real scale
    }
    IS[i] <- prod(S[i,1:expos[i]]) # prob of surviving entire exposure interval
    survive[i] ~ dbern(IS[i]) # does nest survive this nest check interval?
  }
}
", fill = TRUE)
sink()

```


```{r}
parms <- c("dsr.mu")
```


```{r}
# MCMC settings
na <- 1000; ni <- 1100; nt <- 1        
nb <- 100; nc <- 3
```


```{r}
# call jags (~ 4 min)
nest.success_nulo <- jagsUI(dsr.data, inits = NULL, parms, "dsr_null.jags", 
                   n.adapt = na, n.chains = nc, n.thin = nt, 
                   n.iter = ni, n.burnin = nb, parallel = TRUE)
```

```{r}
print(nest.success_nulo, digits = 4)
```



##  Modelo year como predictor


```{r}

sink("dsr_year.jags")
cat("
model {
  # modeling daily nest survival on the logit scale
  dsr.mu ~ dnorm(0, 0.33) # intercept, vague prior (0,1) on real scale
  b.year ~ dnorm(0, 1) # fixed effect for year (vague prior)

  for (i in 1:n.nests) {
    for (j in 1:expos[i]) {
      # daily survival is a function of mean survival and year effect
      logitS[i,j] <- dsr.mu + b.year * year[i] # adding year effect to the logit model
      logit(S[i,j]) <- logitS[i,j] # convert back to real scale
    }
    IS[i] <- prod(S[i,1:expos[i]]) # prob of surviving entire exposure interval
    survive[i] ~ dbern(IS[i]) # does nest survive this nest check interval?
  }
}
", fill = TRUE)
sink()


parms <- c("dsr.mu", "b.year")

# MCMC settings
na <- 1000; ni <- 1100; nt <- 1        
nb <- 100; nc <- 3

# call jags (~ 4 min)
nest.success_year <- jagsUI(dsr.data, inits = NULL, parms, "dsr_year.jags", 
                   n.adapt = na, n.chains = nc, n.thin = nt, 
                   n.iter = ni, n.burnin = nb, parallel = TRUE)
```


```{r}

year = 1:3

dsr <- plogis(nest.success_year$mean$dsr.mu + nest.success_year$mean$b.year * year)

dsr_IC2.5 <- plogis(nest.success_year$q2.5$dsr.mu + nest.success_year$mean$b.year * year)

dsr_IC97.5 <- plogis(nest.success_year$q97.5$dsr.mu + nest.success_year$mean$b.year * year)


df_pre_year <- data.frame(dsr = dsr,
                         IC95L = dsr_IC2.5,
                         IC95H = dsr_IC97.5,
                         year = year)


ggplot(df_pre_year, aes(x= year, y =dsr)) +
  geom_errorbar(aes(ymin= IC95L,
                  ymax= IC95H),
                width = 0.4,
              alpha= 0.4,
              fill= "lightgreen")+
  geom_point()+
  ylim(c(0.9, 1)) +
  theme_classic()
```


##  Modelo year + age como predictor

```{r}
sink("dsr_year_age.jags")
cat("
model {
  # modeling daily nest survival on the logit scale
  dsr.mu ~ dnorm(0, 0.33) # intercept, vague prior (0,1) on real scale
  b.year ~ dnorm(0, 1) # fixed effect for year (vague prior)
  b.age ~ dnorm(0, 1) # fixed effect for age (vague prior)

  for (i in 1:n.nests) {
    for (j in 1:expos[i]) {
      # daily survival is a function of mean survival, year effect, and age effect
      logitS[i,j] <- dsr.mu + b.year * year[i] + b.age * age[i] # adding both year and age effects
      logit(S[i,j]) <- logitS[i,j] # convert back to real scale
    }
    IS[i] <- prod(S[i,1:expos[i]]) # prob of surviving entire exposure interval
    survive[i] ~ dbern(IS[i]) # does nest survive this nest check interval?
  }
}
", fill = TRUE)
sink()


```


```{r}
parms <- c("dsr.mu", "b.year", "b.age")

# MCMC settings
na <- 1000; ni <- 1100; nt <- 1        
nb <- 100; nc <- 3

# call jags (~ 4 min)
nest.success_year_age <- jagsUI(dsr.data, inits = NULL, parms, "dsr_year_age.jags", 
                   n.adapt = na, n.chains = nc, n.thin = nt, 
                   n.iter = ni, n.burnin = nb, parallel = TRUE)
```


```{r}
age <- 1:35

dsr <- plogis(nest.success_year_age$mean$dsr.mu + nest.success_year_age$mean$b.age*age)

dsr_IC2.5 <- plogis(nest.success_year_age$q2.5$dsr.mu + nest.success_year_age$q2.5$b.age*age)

dsr_IC97.5 <- plogis(nest.success_year_age$q97.5$dsr.mu + nest.success_year_age$q97.5$b.age*age)


Age_df_pre <- data.frame(age = age,
                         dsr =dsr,
                         IC95L = dsr_IC2.5,
                         IC95H = dsr_IC97.5)


ggplot(Age_df_pre, aes(x= age, y =dsr)) +
  geom_ribbon(aes(ymin= IC95L,
                  ymax= IC95H),
              alpha= 0.4,
              fill= "lightgreen")+
  geom_line()+
  theme_classic()
```
## Seleccion de modelo

## Evaluacion de modelo


## Crear funcion del modelo

```{r}
# Shaffer's nest survival model, coded for JAGS
{sink("dsr.jags"); cat("
model {
# modeling daily nest survival on the logit scale
dsr.mu ~ dnorm(0, 0.33) # intercept, note this is vague (0,1) on real scale

# fixed effects for nest level covariates (all are linear on logit scale)
b.age ~ dnorm(0, 1) # nest age effect
b.dist ~ dnorm(0, 1) # distance from management edge
b.init ~ dnorm(0, 1) # nest initiation date effect

# note: I'm modeling site and year as crossed random effects - they could also be nested
# random site effects
dsr.area.sd ~ dunif(0.01, 1) # random site effect
dsr.area.tau <- pow(dsr.area.sd, -2)
for (j in 1:n.areas){
  eps.area[j] ~ dnorm(0, dsr.area.tau)
}

# random year effects
dsr.yr.sd ~ dunif(0.01, 1) # random year effect
dsr.yr.tau <- pow(dsr.yr.sd, -2)
for (i in 1:n.yrs){
  eps.yr[i] ~ dnorm(0, dsr.yr.tau)
}

# random species effects
dsr.spp.sd ~ dunif(0.01, 1) # random species effect
dsr.spp.tau <- pow(dsr.spp.sd, -2)
for (i in 1:n.spp){
  eps.spp[i] ~ dnorm(0, dsr.spp.tau)
}

for (i in 1:n.nests){
  for (j in 1:expos[i]){
      # daily survival is function of mean dsr, nest age effect, distance from edge,
      # initiation date, and random effects for area, year, and species
      logitS[i,j] <- dsr.mu + b.age*(age[i]+j-1) + b.dist*dist[i] + b.init*init[i] + 
        eps.area[area[i]] + eps.yr[year[i]] + eps.spp[spp[i]]
      logit(S[i,j]) <- logitS[i,j] # convert back to real scale
      } 
    IS[i] <- prod(S[i,1:expos[i]]) # prob of surviving entire exposure interval
    survive[i] ~ dbern(IS[i]) # does nest survive this nest check interval?
    }
  } # end jags model
",fill = TRUE); sink()}
```




## Definir parametros del modelo

```{r}
parms <- c("dsr.mu", "b.age", "b.dist", "b.init", 
           "dsr.area.sd", "dsr.yr.sd", "dsr.spp.sd",
           "eps.area", "eps.yr", "eps.spp")

```


## Ajustar modelo

```{r}
# MCMC settings
na <- 1000; ni <- 1100; nt <- 1        
nb <- 100; nc <- 3

# call jags (~ 4 min)
# nest.success <- jagsUI(dsr.data, inits = NULL, parms, "dsr.jags", 
#                   n.adapt = na, n.chains = nc, n.thin = nt, 
#                   n.iter = ni, n.burnin = nb, parallel = TRUE)
```


